---
title: Probability Calculations for Means (Practice)
subtitle: Applying the CLT
format:
  html:
    self_contained: true
    embed_resources: true
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
# Ignore this bit of code.  This makes it so that you can render the qmd file even when there are errors in your code.
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```


# Instructions

Answer the following questions, render the document and submit the `.html` report.

# Questions

## GPA

Suppose the mean GPA of BYU-Idaho students is 3.5 and the standard deviation is 0.7. It is well known that this distribution is left-skewed. A random sample of n = 81 students will be drawn.

Use the following R code to answer the questions below:

```{r}
# The specific value
xbar <- 3.3

# The mean of the population
mu <- 3.5

sigma <- 0.7 # The sd of the population

n <- 81 # The random sample amount

sigma_xbar <- sigma / sqrt(n) # The sd of the samples
sigma_xbar

z <- (xbar - mu) / sigma_xbar
z

pnorm(z)

# Area to the left:
pnorm(z)
  
# Area to the right:
1 - pnorm(z)
  
```

__Question__: What is the mean of the distribution of the sample means ($\mu_{\bar{x}}$) for all possible samples of size 81 that could be drawn from the parent population of GPAs?  
__Answer__:  3.5, it stays the same since n > 30

__Question__: What is the standard deviation  of the distribution of the sample means ($\sigma_{\bar{x}}$) for all possible samples of size 81 that could be drawn from the parent population of GPAs?  
__Answer__:  0.7 / sqrt(81) = 0.077777778

__Question__: What is the shape of the distribution of the sample means for all possible samples of size 81 that could be drawn from the parent population of GPAs?  
__Answer__:  symmetric, the bell shape


__Question__: What is the probability that the mean GPA for 81 randomly selected BYU-Idaho students will be less than 3.3?  
__Answer__:  0.005063995


## GRE

Scores on the quantitative portion of the GRE are approximately normally distributed with mean, $\mu = 150.8$ and standard deviation, $\sigma = 8.8$.
 
Use the following R code to answer the questions below:

```{r}
xbar <- 160
mu <- 150.8
sigma <-8.8
n <- 5
sigma_xbar <- sigma / sqrt(n)
sigma_xbar

z <- (xbar - mu) / sigma_xbar
z

# Area to the left:
  
# Area to the right:
1- pnorm(z)

# Percentile (qnorm())
qnorm(.95, 150.8, 8.8)


```

__Question__: Dianne earned a score of 160 on the quantitative portion of the GRE. What is the z-score corresponding to Dianne’s score?  
__Answer__: 1.045455

__Question__: What is the probability that a randomly selected student will score *above* 160 on the quantitative portion of the GRE?  
__Answer__: 0.1479065

__Question__: What GRE score (n=1) corresponds to the 95th percentile?  
__Answer__: 165.2747

__Question__:  What is the probability that the *average* GRE score of 5 randomly selected students will be above 160?  
__Answer__: 0.009701215


## Coast Guard

The United States Coast Guard assumes the mean weight of passengers in commercial boats is 185 pounds. The previous value was lower, but was raised after a tragic boating accident. The standard deviation of passenger weights is 26.7 pounds. The weights of a random sample of 48 commercial boat passengers were recorded. The sample mean was determined to be 177.6 pounds. 

Use the following R code to answer the questions below:

```{r}
xbar <- 177.6
mu <- 185
sigma <- 26.7
n <- 48
sigma_xbar <- 177.6
sigma_xbar

z <- (xbar - mu) / sigma_xbar
z

# Area to the left:
pnorm(z)

pos <- 2 * pnorm(z)
pos
# Area to the right:


```


__Question__: Find the probability that a random sample of passengers will have a mean weight that is as extreme or more extreme (either above or below the mean) than was observed in this sample.  
__Answer__:  0.9667644


__HINT__:  To get a value "as extreme" means it could be higher or lower.  Because the normal distribution is symmetric, if the $z<0$, we can take the area to the *left* of $z$ and multiply by 2.  Conversely, if $z>0$ we can take the area to the right and multiply by 2.  


## Tankers

Tanker trucks are designed to carry huge quantities of gasoline from refineries to filling stations. A factory that manufactures the tank of the trucks claims to manufacture tanks with a capacity of 8550 gallons of gasoline. The actual capacity of the tanks is normally distributed with mean, $\mu = 8544$ gallons, and standard deviation, $\sigma=12$ gallons.

Use the following R code to answer the questions below:

```{r}
xbar <- 8550
mu <- 8544
sigma <- 12
n <- 20
sigma_xbar <- sigma / sqrt(n)
sigma_xbar

z <- (xbar - mu) / sigma_xbar
z

# Area to the left:
pnorm(z)

#Area to the right
1-pnorm(z)

## Area between A and B
A <- 8541
B <- 8547

pnorm(B, mu, sigma_xbar) - pnorm(A, mu, sigma_xbar)

```


__Question__: Find the **z-score** corresponding to a single tank ($n=1$) with a capacity of 8550 gallons. Round your answer to one decimal place.  
__Answer__: 0.5

__Question__: What is the probability that a randomly selected tank will have a capacity of **less** than 8550 gallons?  
__Answer__:  0.6914625

__Question__: A simple random sample of $n = 20$ tanks was selected. Find the z-score corresponding to a sample mean capacity for 20 tanks of 8550.  
__Answer__: 2.236068

__Question__: What is the probability that the sample mean of $n=20$ randomly selected tanks will be between 8541 and 8547?  
__Answer__: 0.7364475


