---
title: Intro to Data Wrangling
format:
  html:
    self_contained: true
    embed_resources: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```
<!---
-->

The 5 steps of conducting the data and study:
| Step 1: | **D**aniel   | **D**esign the study   |
| Step 2: | **C**an      | **C**ollect data       |
| Step 3: | **D**iscern  | **D**escribe the data  |
| Step 4: | **M**ore     | **M**ake inferences    |
| Step 5: | **T**ruth    | **T**ake action        |

-----------------------------------------------------------------
Groups in the experiment:
considered group: The group which has the key factor influence in it.
control group: The group which don't have the key factor influence.

-----------------------------------------------------------------
Data Type: 
Quantitative Data: The data can be measurable and mathematically operated. Ex. Age
Categorical Data: The data can't be measured and mathematically operated. Ex. Birth Month

-----------------------------------------------------------------
Study Type:
Observational Studies: The study that only observe, not controlling any factors and change the object artificially.
Experiments: The study that you control several factors to have different results.

-----------------------------------------------------------------
Experiments objects:
Experimental units or subjects: The objects(people/ items/ animals.etc) that receive the experiments. 
                                Ex. The mouse which receive the test.
Treatments: The key factor. Ex.The vaccine.
Population: The whole population body which is relevant to the experimental units/ subjects.

-----------------------------------------------------------------
Sampling scheme:

- simple random sample: Pick the samples randomly. I mean, randomly for real, like a blind pick.

- stratified sample: Separate the data into groups by their traits(Ex. school year, sex, race...etc)

- Systematic sample: Pick the sorted (or arrayed) data by the same difference value between each sample. 
                     (Ex. The difference is 4. Then pick the 1st, 5th, 9th, 13th data from the dataset)
- cluster/block sample: Separate the whole data set into chunks/blocks , then When the variation from one block to another is relatively low, compared to the variation within the block.

-----------------------------------------------------------------
Short Keys: 
Ctrl + Alt + i : Create a R language chuck 

```{r}

```










