---
title: Don't Take it Personally
subtitle: Hypothesis Tests and Confidence Intervals
format:
  html:
    self_contained: true
    embed_resources: true
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```



# Introduction

In this activity, you will execute statistical hypothesis tests and generate confidence intervals for each of the Big 5 personality traits using data collected from a random sample of Brother Cannon's Math 221 students.  

__Question__:  What is the population of this analysis?  
__Answer__:  Brother Cannon's all MATH 221D students

For each personality trait, include:

1. A statement of the null and alternative hypotheses and why you chose the alternative you did.
2. Choose alpha, $\alpha = $
3. Check that you can trust the normality of the mean (`n > 30` or `qqPlot(respons_variable)`)
4. Run the one-sample t-test and state your conclusion (technical and contextual explanation)
5. Calculate a $1-\alpha$ level confidence interval and describe in words what it means in context of the research question

Recall that we can use `favstats()` to get summary statistics, `boxplot()` and `histogram()` to get visualizations, and the `t.test()` function to get hypothesis tests and confidence intervals.  **Be sure to label your plots' axes and include a title.**  

```{r}

# Load Libraries

library(tidyverse)
library(mosaic)
library(rio)
library(car)


# Load Data
big5 <- import('https://raw.githubusercontent.com/byuistats/Math221D_Cannon/master/Data/All_class_combined_personality_data.csv')
glimpse(big5)
```


# Extroversion

State your null and alternative hypotheses:  

$$H_o:  \mu = 50$$  
$$H_a:  \mu > 50$$  

$$\alpha = 0.025 $$  
1. Create a table of summary statistics:  

```{r}
# Extroversion

favstats(big5$Extroversion) %>% knitr::kable()
```

2. Create a histogram of Extroversion:  

```{r}
### I'm making a single variable called "extrov" that drops the missing values.  You can do the same thing with the other traits to make analysis a little easier
extrov <- na.omit(big5$Extroversion) 

histogram(extrov, xlab = "Extroversion", main = "Histogram of Extroversion Percentiles")
```

3. Perform the one-sample t.test:  

```{r}
# For Hypothesis Test:
t.test(extrov, mu = 50, alternative = "greater")
```

4. Explain your conclusion:  

Technical:  Because the p-value is less than $\alpha$, I reject the null hypothesis.  

Contextual:  I have sufficient evidence to suggest that Brother Cannon's students are, on average, more Extroverted than the general population.  

5. Create a Confidence Interval for the average extroversion of Brother Cannon's students:  

```{r}
# For a Confidence Interval:
t.test(extrov, conf.level = 1-.025)$conf.int
```

6. Explain your confidence interval:  

I am 97.5% confident that the true average extroversion of Brother Cannon's students is somewhere between the 54.62 and 59.34 percentiles.  

# Agreeableness

State your null and alternative hypotheses.  For example, do you think Brother Cannon's students are more, less, or just different than the general population?  

$$H_o:  \mu = 60$$  
$$H_a:  \mu > 60$$  

$$\alpha =  0.025$$    
1. Create a table of summary statistics for Agreeableness:  

```{r}
favstats(big5$Agreeableness) %>% knitr::kable()
```

2. Create a histogram of Agreeableness:  

```{r}
agreea <- na.omit(big5$Agreeableness)

histogram(agreea, xlab = "Agreeableness", main = "Histogram of Agreeableness Percentiles")
```

3. Perform the one-sample t.test:  

```{r}
t.test(agreea, mu = 60, alternative = "greater")
```

4. Explain your conclusion:  

Technical:  Because the P-value is less than alpha value (0.025), I reject the null hypothesis.

Contextual:  I have sufficient evidences to suggest that the average agreeableness score of Brother Cannon's students is more than 60.

5. Create a Confidence Interval for the average agreeableness of Brother Cannon's students:  

```{r}
# For a Confidence Interval:
t.test(agreea, conf.level = 1-0.025)$conf.int
```

6. Explain your confidence interval:  
I'm 97.5% sure that the real agreeableness mean score of Brother Cannon's students is between 71.95341 and 74.91572 percentiles.



# Openness
State your null and alternative hypotheses:  

$$H_o:  \mu = 75$$  
$$H_a:  \mu ≠ 75$$  

$$\alpha = 0.025$$  

1. Create a table of summary statistics for Openness:  

```{r}
favstats(big5$Openness) %>% knitr::kable()
```

2. Create a histogram of Openness:  

```{r}
openn <- na.omit(big5$Openness)

histogram(openn, xlab = "Openness", main = "Histogram of Openness Percentiles")
```

3. Perform the one-sample t.test:  

```{r}
t.test(openn, mu = 75, alternative = "two.sided")
#You can ignore the alternative for two.sided since it's default
```

4. Explain your conclusion:  

Technical:  Since the P-value(0.0001075) is less than alpha value(0.025), I reject the hypothesis.

Contextual:  We have sufficient evidence to suggest the true average Openness percentiles of Brother Cannon's students is not equal to 75.

5. Create a Confidence Interval for the average openness of Brother Cannon's students:  

```{r}
# For a Confidence Interval:
t.test(openn, conf.level = 1-0.025)$conf.int
```

6. Explain your confidence interval:  
I'm 97.5% confident that the real average openness percentile of Brother Cannon's students is located between 69.97700 and 73.64523 percentiles.


# Neuroticism

State your null and alternative hypotheses:  

$$H_o:  \mu = 50$$  
$$H_a:  \mu > 50$$  

$$\alpha =  0.025$$  
1. Create a table of summary statistics for Neuroticism:  

```{r}
favstats(big5$Neuroticism) %>% knitr::kable()
```

2. Create a histogram of Neuroticism:  

```{r}
neurot <- na.omit(big5$Neuroticism)

histogram(neurot, xlab = "Neuroticism", main = "Histogram of Neuroticism Percentile")
```

3. Perform the one-sample t.test:  

```{r}
t.test(neurot, mu = 50, alternative = "greater")
```

4. Explain your conclusion:  

Technical:  Since the P-value (0.998) is bigger than alpha value, I won't reject the hypothesis.

Contextual:  I have sufficient evidences to suggest that the average of Neuroticism percentiles of Brother Cannon's students is 50.

5. Create a Confidence Interval for the average neuroticism of Brother Cannon's students:  

```{r}
# For a Confidence Interval:
t.test(neurot, conf.level = 1-0.025)$conf.int
```

6. Explain your confidence interval:  
I'm 97.5% confident that the true average of Neuroticism percentiles of Brother Cannon's student is located between 45.35754 and 49.42270 percentiles.


# Conscientiousness

State your null and alternative hypotheses:  

$$H_o:  \mu = 50$$  
$$H_a:  \mu > 50$$  

$$\alpha = 0.025$$  
1. Create a table of summary statistics:  

```{r}
favstats(big5$Conscientiousness) %>% knitr::kable()
```

2. Create a histogram of Conscientiousness:  

```{r}
cons <- na.omit(big5$Conscientiousness)

histogram(cons, xlab = "Conscientiousness", main = "The Histogram of Conscientiousness")
```

3. Perform the one-sample t.test:  

```{r}
t.test(cons, mu = 50, alternative = "greater")
```

4. Explain your conclusion:  

Technical:  Since the p-value(< 2.2e-16) is less than alpha(0.025), I reject the hypothesis.

Contextual: I have sufficient evidence to suggest the real average of Conscientiousness Percentiles of Brother Cannon's students is greater than 50.
# Bring the conclusion back to the study itself

5. Create a Confidence Interval for the average conscientiousness of Brother Cannon's students:  

```{r}
# For a Confidence Interval:
t.test(cons, conf.level = 1-0.025)$conf.int
```

6. Explain your confidence interval:  
I'm 97.5% confident that the real average of Conscientiousness Percentiles of Brother Cannon's students is located between 59.22813 and 62.92989


