---
title: 2-Sample Independent T-Test Practice

format:
  html:
    self_contained: true
    embed_resources: true
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```



# Instructions

Here are several opportunities to practice analyzing 2-sample independent t-tests using R.  For each question, you will:

- Read in data
- Identify the Response and Independent variable
- Create data summaries (numerical and graphical)
- Statistically analyze the data
- Check for the suitability of the statistical test (CLT, Normality)
- State your hypothesis test conclusions and interpret your confidence intervals

When you finish, render this document and submit the `.html` in Canvas.  

```{r}
# Load the libraries

library(rio)
library(mosaic)
library(tidyverse)
library(car)

```


# It's a Date!

Dating Behavior was studied in a speed dating experiment where random matching was generated and created randomization in the number of potential dates.  In a survey to each participant, one question was asked about how important attraction is when they date. The attraction value is on a scale from 0 (not important at all) to 100 (very important).  

The researchers want to determine if males report to value attractiveness more than females.  Use a level of significance of 0.05.

## Load the Data

```{r}
dating <- read_csv('https://github.com/byuistats/Math221D_Course/raw/main/Data/dating_attractive_longformat.csv')
view(dating)

```

## Explore the Data

__Question__:  What is the *response* variable?  
__Answer__:  Importance score

__Question__:  What is the *explanatory* variable?  
__Answer__:  Biosex

Create a side-by-side boxplot for the amount of reported importance of attractiveness for each biosex.

Add a title and change the colors of the boxes.

```{r}
boxplot(dating$Importance ~  dating$Biosex, main = "Attraction Importance of Dating between Genders", col = c(2,3))

```

What do you observe?  

Create a table of summary statistics for each group (`favstats()`):  


```{r}
favstats(dating$Importance ~  dating$Biosex)
```


## Hypothesis Test

State your null and alternative hypotheses (replace the ??? with the appropriate symbol):  

$$H_0:  \mu_{F} = \mu_{M}$$  

$$H_a:\mu_{F} < \mu_{M}$$

__NOTE__: The default for R is to set group order alphabetically.  This means Group 1 = Female.


Check that the samples for both groups are normally distributed
```{r}
qqPlot(dating$Importance~dating$Biosex)
```

Do the data for each group appear normally distributed? Yes, since the data number is big enough(>30).

Why is it OK to continue with the analysis?
Yes.

Perform a t-test.

```{r}
t.test(dating$Importance ~ dating$Biosex, mu = 0, alternative = "less") #mu = 0 is the default

```

__Question__:  What is the value of the test statistic?  
__Answer__:  t = -8.9016

__Question__:  How many degrees of freedom for this test?  
__Answer__:  df = 495.36

__Question__:  What is the p-value?  
__Answer__:  < 2.2e-16

__Question__:  What do you conclude?  
__Answer__:  The p-value is small, reject the Ho. I have sufficient evidence to suggest true difference in means of attraction in dating between group females and group males is less than 0.

### Confidence Interval

Create a confidence interval for the difference of the average Importance Score between both groups:

```{r}
t.test(dating$Importance ~ dating$Biosex, conf.level = 0.975)$conf.int

```
Conclusion: 
I'm 97.5% confident that the true difference in means of the importance of attraction in dating between group female and group male is located between -11.559206 and -6.897651.


# Tooth hurty

A simple random sample of dental bill costs was collected at offices in Rexburg and Idaho Falls.  Let $\mu_{IF}$ be the *population* mean of dental costs in Idaho Falls and $\mu_{R}$ be the *population* mean of dentals costs in Rexburg.  

We suspect that dental costs are higher in Rexburg because of less competition.  

Use the data imported below to answer the following questions.

```{r}
dentist <- read_csv('https://github.com/byuistats/Math221D_Course/raw/main/Data/DentistOfficeBills_longformat.csv')

glimpse(dentist)
```

## Review The Data

__Question__:  What is the *response* variable?  
__Answer__:  The mean of dental cost

__Question__:  What is the *explanatory* variable?  
__Answer__:  The location


Create summary statistics tables of dental costs for each office:

```{r}
favstats(dentist$Cost ~ dentist$Location) %>% knitr::kable()
```

Create a side-by-side boxplot for dental costs for each office.

```{r}
boxplot(dentist$Cost ~ dentist$Location, main = "Average Dental Costs in Idaho Falls and Rexburg",col =  c(2,3))
```

Check the normality of each group.

```{r}
qqPlot(dentist$Cost ~ dentist$Location)
```

__Question__: Do the samples from both groups appear to be normally distributed?  If not, is it a cause for concern for our statistical inference? Yes it is normal distributed.
A: According to QQPlot graph, both groups are normal distributed.

## Hypothesis Test

State your null and alternative hypotheses (replace the question marks with the appropriate symbols):

$$H_0:  \mu_{IF} = \mu_{R}$$  

$$H_a:\mu_{IF} < \mu_{R}$$

Perform the appropriate analysis:

```{r}
t.test(dentist$Cost ~ dentist$Location, alternative = "less")
```


__Question__:  What is the test statistic?  
__Answer__: t = -0.43437

__Question__:  What is the P-value?  
__Answer__:  p-value = 0.3332

State your conclusion:  Since the p-value is big enough(0.3332 > 0.05), I have sufficient evidence to suggest than the true difference of dental cost between Idaho Falls(260.059) is less than Rexburg(288.597).



## Confidence Interval

Create a confidence interval for the difference in costs between the IF and Rexburg offices:

```{r}
t.test(dentist$Cost ~ dentist$Location, conf.level = 0.95)$conf.int
```

Explain the confidence interval in context of the research question: 
I'm 95% confident that the true difference of dental cost in means between Idaho Falls and Rexburg is located between -161.5398 and 104.4638.



#  Birth Weights

A study was conducted in which researchers obtained birth weights of infants born in Illinois during the same year. The birth weights are categorized based on the race and birthplace of the mother. Researchers want to know if there is a difference in mean birth weights of babies born to black women who were born in the United States and babies born to black women who were born in Africa. Values were recorded (in grams) for each infant born. The data are recorded in the file. 

Let $\mu_A$ be the *population* average of babies born to African-born mothers and $\mu_{IL}$ be the *population* mean weight of babies born to mothers in Illinois.


Use the data imported below to answer the following questions.

```{r}
birth_weights <- read_csv('https://github.com/byuistats/Math221D_Course/raw/main/Data/IllinoisBirthWeightsTwoVar_longformat.csv')

glimpse(birth_weights)
```

## Review The Data

__Question__:  What is the *response* variable?  
__Answer__:  weight_g, which is the weight of new born

__Question__:  What is the *explanatory* variable?  
__Answer__:  Birth_Place, which is the birth place of the mother of the new born


Create summary statistics tables of birth weights for each country:

```{r}
favstats(birth_weights$Weight_g ~ birth_weights$Birth_Place) %>% knitr::kable()
```

Create a side-by-side boxplot for birth weights for each country:

```{r}
boxplot(birth_weights$Weight_g ~ birth_weights$Birth_Place, main = "The Weigths of New Borns from Mothers from Different Birthplace" ,col = c(2,3))
```

Check the normality of each group.

```{r}
qqPlot(birth_weights$Weight_g ~ birth_weights$Birth_Place)
```

__Question__: Do the samples from both groups appear to be normally distributed?  If not, is it a cause for concern for our statistical inference?
: According to the QQPlot and number of data, both of the groups are normal distributed.

## Hypothesis Test

State your null and alternative hypotheses (replace the question marks with the appropriate symbols):

$$H_0:  \mu_{A} = \mu_{IL}$$  

$$H_a:\mu_{A} > \mu_{IL}$$

Perform the appropriate analysis:

```{r}
t.test(birth_weights$Weight_g ~ birth_weights$Birth_Place, alternative = "greater")
```


__Question__:  What is the test statistic?  
__Answer__: t = 1.9021

__Question__:  What is the P-value?  
__Answer__: p-value = 0.0306

State your conclusion:  Since the p-value is small (0.0306 < 0.05), I have sufficient evidence to suggest the newborns from the African-born mothers is heavier than the the newborns from the Illinois-born mothers.

## Confidence Interval

Create a confidence interval for the average difference in weights between babies born to mothers in Africa and Illinois:

```{r}
t.test(birth_weights$Weight_g ~ birth_weights$Birth_Place, conf.level = 0.95)$conf.int
```

Explain the confidence interval in context of the research question:  
I'm 95% confident to suggest the average weight difference between newborns from Africa-born mothers is -11.76601 to 499.76455 grams heavier than the newborns from Illinois-born mothers.

#  Miscellany

__Question__:  A financial economist is studying married couples in which both spouses work. He wants to compare the mean income earned by husbands with the mean income earned by their wives. Should he use independent sampling or dependent sampling, and why?
__Answer__:  He should use dependent sampling since husband and wife are related. For one group, if you choose husband, than another group, or the corresponding sample would be his wife. Vice versa.









