---
title: ANOVA Practice
format:
  html:
    self_contained: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

# Life Expectancy of Female Authors

## Introduction

You are curious to compare life expectancy between female poets, novelists, and non-fiction writers.  

You take a sample of female authors from each of the three groups to test if the average age at death is different between any of the three types of authors using a level of significance of, $\alpha = 0.05$.

## Load the Data and Libraries


```{r}
library(rio)
library(tidyverse)
library(mosaic)
library(car)

womenpoet <- rio::import("https://byuistats.github.io/BYUI_M221_Book/Data/womenpoet.xls")
glimpse(womenpoet)
```

## Explore the Data

Create a side-by-side boxplot of the age at death of each of the different author styles.

Modify the colors of each of the boxes for each group.

```{r}
boxplot(womenpoet$Age ~ womenpoet$Type, main = "The Age of Different Women Writers", col = c(2,3,4))
```


Create a summary statistics table for age at death for each author type:

```{r}
favstats(womenpoet$Age ~ womenpoet$Type) %>% knitr::kable()

```

List the mean and standard deviations of age at death for:

|womenpoet$Type |     mean|       sd|  n| missing|
|:--------------|--------:|--------:|--:|-------:|
|Nonfiction     | 76.87500| 14.09691| 24|       0|
|Novels         | 71.44776| 13.05151| 67|       0|
|Poems          | 63.18750| 17.29710| 32|       0|

## Perform the Appropriate Analysis

State your null and alternative hypotheses:

$$ H_o = u_{Nonfiction} = u_{Novel} = u_{Poem}$$

$$ H_a = At\ least\ one\ of\ the\ u\ not\ equals\ to\ another\ one. $$

Perform an Analysis of Variance test including checking for the appropriateness of our analysis.

```{r}
aov_women_writer <- aov(womenpoet$Age ~ womenpoet$Type)
aov_women_writer

summary(aov_women_writer)

```

__Question__:  What is the test statistic (F-value)?
__Answer__:  6.563


__Question__:  What are the degrees of freedom for your analysis?
  a. Numerator (between Groups) Degrees of Freedom
  b. Denominator (within groups) Degrees of Freedom
__Answer__:  a = 2; b = 120


__Question__:  What is the P-value?
__Answer__:  0.00197


__Question__:  Do you reject the null hypothesis?  Why?
__Answer__:  Yes, I will reject the null hypothesis since the P-value is less than alpha-value (0.00197 < 0.05)


__Question__:  State your conclusion in context of the problem.
__Answer__:  I have sufficient evidence to suggest that the average age among women writers in Nonficition, Novel, and Poem field are not all equal to each other.

```{r}
sds = favstats(womenpoet$Age ~ womenpoet$Type)$sd

max(sds) / min(sds)

qqPlot(aov_women_writer$residuals)
```

__Question__:  Can we trust the p-value?
  a. Check for equal standard deviation (is the ratio of the largest SD / smallest SD greater than 2?)
  b. Check Normality of the residuals (`qqPlot()`)
__Answer__:  Yes. According to the result of equal standard deviation test (1.325295 < 2, test passed) and qqPlot graph (graph shows that the residuals are normal distributed), we can trust the p-value. 


# Testing Mosquitos

## Introduction

A study was conducted to determine if different types of material can reduce the amount of mosquito human contact.  The researchers evaluated five different types of patches 1=Odomos, 2=Deltamethrin, 3=Cyfluthrin, 4=D+O, 5=C+O.  

The amount of mosquito human contact was measured to assess any differences between the five different types of material. Use a level of significance of 0.05.

## Load the Data

```{r}

MosquitoPatch <- rio::import("https://raw.githubusercontent.com/rdcromar/Math221D/main/MosquitoPatch.csv") %>% mutate(Treatment = factor(Treatment))

glimpse(MosquitoPatch)
```


## Review the Data

Create a side-by-side boxplot for human contact for each of the treatment groups.

Add a title and change the colors of the boxes.  

```{r}
boxplot(MosquitoPatch$mosqs ~ MosquitoPatch$Treatment, main = "The Amount of Mosquito Human Contact with Different Patches", col = c(2,3,4,5,6))

```

Create a summary statistics table for human contact for each of the treatment groups:

```{r}
favstats(MosquitoPatch$mosqs ~ MosquitoPatch$Treatment) %>% knitr::kable()

```

__Question__:  What do you observe?
__Answer__:  The patch which has the best performance in average is the Deltamethrin, the number 2 in the treatment column.

__Question__:  What is the maximum standard deviation?
__Answer__:  18.01, from patch Odomos

__Question__:  What is the minimum standard deviation?
__Answer__:  14.10 ,from patch Odomos + Cyfluthrin & Odomos + Deltamethrin


## Perform the Appropriate Analysis

State your null and alternative hypotheses:

$$ H_o = u_{Odomos} = u_{Deltamethrin} = u_{Cyfluthrin} = u_{D+O} = u_{C+O}$$

$$ H_a = At\ least\ one\ of\ the\ u\ not\ equals\ to\ another\ one. $$

Perform an Analysis of Variance test including checking for the appropriateness of our analysis.

```{r}
aov_MosPatches <- aov(MosquitoPatch$mosqs ~ MosquitoPatch$Treatment)
aov_MosPatches

summary(aov_MosPatches)
```

__Question__:  What is the test statistic (F-value)?
__Answer__:  4.48

__Question__:  What are the degrees of freedom for your analysis?

  a. Numerator (between Groups) Degrees of Freedom
  b. Denominator (within groups) Degrees of Freedom
__Answer__:  a = 4; b = 145
  
__Question__:  What is the P-value?
__Answer__:  0.00192

__Question__:  Do you reject the null hypothesis?  Why?
__Answer__:  Yes, since the P-value is less than alpha-value (0.00192 < 0.05)

__Question__:  State your conclusion in context of the problem.
__Answer__:  I have sufficient evidence to suggest the average performance of reducing the amount of mosquito human contact among the 5 patches not all equal to each others.


```{r}
sds = favstats(MosquitoPatch$mosqs ~ MosquitoPatch$Treatment)$sd
max(sds) / min(sds)

qqPlot(aov_MosPatches$residuals)
```

__Question__:  Can we trust the p-value?
  a. Check for equal standard deviation (is the ratio of the largest SD / smallest SD greater than 2?)
  b. Check Normality of the residuals (`qqPlot()`)
__Answer__:  Yes. According to the equal standard deviation test (1.149441 < 2, test passed) and the qqPlot graph of residuals (The dots are distributed in the blue area, which means the residuals are normal distributed), we can trust the P-value.




